Significant wind disturbances induced by giant dunes.

### 2021-12-10 Paper/main.tex
Rewriting section 3.2

  Another way to increase the flow deflection is its confinement below a capping surface, that result in further streamline compression. This happens when the flow disturbance induced by the obstacle reaches the surface. As obstacles typically disturb flow over a characteristic height similar to their width, this is well captured by the parameter $k H$, where $k = 2\pi/\lambda$ is the wavenumber and $H$ the ABL depth. Here, the giant dunes have kilometric wavelengths, such that $0.02 \lesssim k H \lesssim 5$, and they interact most of the time with the capping layer and the stratified free atmosphere above \citep{andreotti2009}.

@@

  The smallest wind disturbances are expected during the day, when the ABL depth is comparable to the dune wavelength ($k H \gtrsim 1$) and for large wind velocities, which correspond to a weak confinement situation (Fig.~\ref{Fig4b}). On the contrary, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by shallow ABL (Fig.~\ref{Fig4d--f}). Note that this strong confinement can be somewhat reduced in the case of strong winds (corresponding to large Froude numbers, see Fig.~\ref{Fig4f}), explaining the threshold in velocity observed in the data (see section~\ref{section_data_feedback}).

@@

  In the spirit of \citet{Sheridan2006}, we aim to compute flow regime diagrams in the space defined by the three relevant non-dimensional numbers presented above, $\left(k H,\, \mathcal{F}r_{\textup{S}}, \, \mathcal{F}r_{\textup{I}}\right)$. They are calculated from the time series of the geopotential, temperature and specific humidity vertical profiles available in the ERA5 climate reanalysis (see SI~\ref{?}). The relative velocity modulation is computed as
  %
  \begin{equation}
    \delta_{\textup{u}} = \frac{u_{*}^{\textup{ERA}} -  u_{*}^{\textup{station}}}{u_{*}^{\textup{ERA}}},
  \end{equation}
  %
  and the flow deviation as the minimal angle between the wind orientation in the two datasets:
  %
  \begin{equation}
    \delta_{\theta} = \left\vert \textup{min}\left(\left[\theta_{\textup{ERA}} - \theta_{\textup{station}}\right] \, \textup{mod.} \, 360, \left[\theta_{\textup{station}} - \theta_{\textup{ERA}}\right] \, \textup{mod.} \, 360\right) \right\vert,
  \end{equation}
  %

  When these two variables are represented in the marginal spaces $(k H,\, \mathcal{F}r_{\textup{S}})$ and $(k H,\, \mathcal{F}r_{\textup{I}})$, different regime emerges (Fig.~\ref{Fig5}). The small wind disturbances ($\delta_{\theta} \to 0$, $\delta_{\textup{u}} \to 0$) are located in the top-right part of the diagrams, corresponding to a regime mixing low-interactions ($k H$ large enough, Fig.~\ref{Fig4}b) and low-confinement ($\mathcal{F}r_{\textup{S}}, \, \mathcal{F}r_{\textup{I}}$ large enough, Fig.~\ref{Fig4}c).

  Lower values of $k H$ (stronger interaction) or Froude numbers (stronger confinement) then both leads to an increase in wind disturbances, both in terms of orientation and velocity. Interestingly, the limit of no-interactions between the topography and the boundary layer structure ($k H \gg 1$), in which the properties of the capping layer and the stratification become irrelevant (Fig.~\ref{Fig4}b--\ref{Fig4}c), is never reached here, in the case of giant dunes.
  %
  Below a threshold value of $k H \simeq 0.3$, wind disturbance occurs independently of the Froude numbers value. However, the latter seem to control a transition between from damped to amplified wind velocities within the interdune (Fig.~\ref{Fig5}c--\ref{Fig5}d), for which we do not have an explanation.

>>>>

  Another way to increase the flow deflection is its confinement below a capping surface, that results in further streamline compression. This happens when the flow disturbance induced by the obstacle reaches the surface. As obstacles typically disturb flow over a characteristic height similar to their width, the potential of interaction between the dunes and the overlying atmospheric structure is well captured by the parameter $k H$, where $k = 2\pi/\lambda$ is the wavenumber and $H$ the ABL depth. Here, the giant dunes have kilometric wavelengths, such that $0.02 \lesssim k H \lesssim 5$, and they interact most of the time with the capping layer and the stratified free atmosphere above \citep{andreotti2009}.

@@

  The smallest wind disturbances are expected during the day, when the ABL depth is comparable to the dune wavelength ($k H \gtrsim 1$) and for large wind velocities, which correspond to a weak confinement situation (Fig.~\ref{Fig4}b). On the contrary, large wind disturbances are expected to occur during the night, when the confinement is mainly induced by shallow ABL (Fig.~\ref{Fig4}d--f). Note that this strong confinement can be somewhat reduced in the case of strong winds (corresponding to large Froude numbers, see Fig.~\ref{Fig4}f), explaining the threshold in velocity observed in the data (see section~\ref{section_data_feedback}).

@@

  To highlight these different regimes from our data, we compute wind disturbance diagrams in the space defined by the three relevant non-dimensional numbers presented above, $\left(k H,\, \mathcal{F}r_{\textup{S}}, \, \mathcal{F}r_{\textup{I}}\right)$. Those are calculated from the time series of the geopotential, temperature and specific humidity vertical profiles available in the ERA5 climate reanalysis (see SI section 2).
  %
  Flow deviation is computed as the minimal angle between the wind orientations from the two datasets:
  \begin{equation}
    \delta_{\theta} = \left\vert \textup{min}\left(\left[\theta_{\textup{ERA}} - \theta_{\textup{station}}\right] \, \textup{mod.} \, 360, \left[\theta_{\textup{station}} - \theta_{\textup{ERA}}\right] \, \textup{mod.} \, 360\right) \right\vert.
  \end{equation}
  %
  The relative velocity modulation is computed as
  %
  \begin{equation}
    \delta_{\textup{u}} = \frac{u_{*}^{\textup{ERA}} -  u_{*}^{\textup{station}}}{u_{*}^{\textup{ERA}}}.
  \end{equation}

  As both Froude numbers have qualitatively the same impact on the flow confinement, we first leave aside the internal Froude number, and focus on the space $(k H,\, \mathcal{F}r_{\textup{S}})$. When representing the two variables $\delta_{\theta}$ and $\delta_{\textup{u}}$ in this space, different regime emerges (Fig.~\ref{Fig5}). Small wind disturbances ($\delta_{\theta} \to 0$, $\delta_{\textup{u}} \to 0$) are located in the top-right part of the diagrams, corresponding to a regime mixing low-interaction and low-confinement ($k H$ and $\mathcal{F}r_{\textup{S}}$ large enough, Fig.~\ref{Fig4}c).
  %
  Lower values of $k H$ (stronger interaction) or Froude numbers (stronger confinement) then both leads to an increase in wind disturbances, both in terms of orientation and velocity.   Below a threshold value of $k H \simeq 0.3$, wind disturbance occurs independently of the Froude numbers value. Furthermore, this also seems to control a transition between from damped to amplified wind velocities within the interdune (Fig.~\ref{Fig5}b), for which we do not have an explanation. Note that the same interpretation can be done with the diagrams including the internal Froude number $\mathcal{F}r_{\textup{I}}$, as shown by Fig.~\ref{Fig12_supp}.

  Interestingly, the limit of no-interactions between the topography and the boundary layer structure ($k H \gg 1$), in which the properties of the capping layer and the stratification become irrelevant (Fig.~\ref{Fig4}b--\ref{Fig4}c), is never reached here, in the case of giant dunes.
